The Effects of Stellar Binarity on Transit Survey Occurrence Rates

### 2017-10-05 paper/ms.tex
wordsmithing

An astronomer who does not believe in binaries wants to measure the
occurrence of planets of a certain type around stars of a certain type.
In other words, they wish to find the mean number of planets within specific 
planetary $(R_p, P)$ bounds orbiting the stars in a given volume 
of stellar $(M_\star, R_\star, L_\star)$ phase-space.
They perform a signal-to-noise limited transit survey and detect $N_{\rm det}$ 
transit signals that appear to be from planets of the desired type.  They 
calculate the number of stars $N_\star$ that appear to be searchable for the 
desired type of planet, and also of the desired stellar type.  These 
``searchable stars'' are the points on the sky for which they think planets 
are observed with 100\% detection efficiency. Correcting for the geometric 
transit probability $f_{\rm g}$, they compute an apparent occurrence rate 
$\Gamma_{\rm a}$:

@@

Here we focus on problems that arise from the fact that stars of the desired
type often exist in binaries. 
We assume for simplicity that all binaries in the transit survey are 
spatially unresolved.

An immediate complication is that~--~due to dynamical stability or some 
aspect of planet formation~--~ the occurrence rate of planets may differ 
between binary and single-star systems.
If ``occurrence rate'' is defined purely as the mean number of planets within 
set radius and period bounds per star in a given interval $\Delta M_\star, 
\Delta R_\star, \Delta L_\star$, it must include an implicit marginalization 
over stellar multiplicity.
Otherwise, one must discuss ``occurrence rates in single star systems'', 
``occurrence rates about primaries of double systems'', and
``occurrence rates about secondaries of double star systems''.

Outside of bonafide astrophysical differences, there are observational biases.
A given apparently-searchable star may truly be a single star of the desired 
type. If not, and ignoring higher order multiples, one of the following must 
be true:
\begin{enumerate}	
\item The apparently-searchable star is a binary system, with two searchable 
stars of the desired type. $N_\star$ is under-counted by one for every such 
system.
%
\item The apparently-searchable star is a binary system, with one searchable 
star of the desired type. 
The searchable star of the desired type could be the primary or the secondary.
$N_\star$ is correctly counted for every such system, though there may be 
systematic errors in estimates of stellar parameters.
%
\item  The apparently-searchable star is a binary system, in which neither of 
the components is a searchable star, but at least one is of the desired type. 
$N_\star$ is over-counted by one for every such system.
\end{enumerate}

@@

The above enumerates possible errors when counting $N_\star$.
When tallying $N_{\rm det}$, the detected signals from planets of the desired 
type, neglecting binarity introduces the following error cases:
\begin{enumerate}
\item 
A detected signal is from a star of undesired type (by the above assumptions, 
the secondary of a binary system), from a planet of undesired type, and is 
incorrectly counted as a planet of desired type.
%
\item
A detected signal is from a star of desired type (the primary or secondary of 
a binary systems), from a planet of undesired type, and is incorrectly counted 
as a planet of desired type.
This and the case above occur when the host star parameters are miscalculated 
({\it e.g.,} the host star is the faint secondary, but is assumed to be 
primary), or when the constant diluting light from the binary companion is not 
corrected.
$N_{\rm det}$ should be lowered by 1 for every such case.
%
\item
A detected signal is from a star of desired type, from a planet of desired 
type, but incorrectly counted as planet of undesired type.
$N_{\rm det}$ should be raised by 1 for every such case.
\item
An {\it undetected} signal from a planet of the desired type, around a star of 
the desired type, was not detected because of dilution from the companion 
(because the SNR floor of the survey is set for single stars). 
This happens during case \#3 in $N_\star$-counting errors (the 
apparently-searchable star is a binary system, in which neither component is 
in fact searchable). If $N_\star$ is corrected for this, {\it i.e.,} the 
system is correctly counted as ``not searchable'', no correction to $N_{\rm 
det}$ is needed.
\end{enumerate}

@@

Our approach is step-by-step starting from a very simple scenario.
In Sec.~\ref{sec:model_1} we first consider a universe consisting of only the 
desired types of stars and planets.  
We give an analytic expression for $\Gamma_{\rm a}/\Gamma_{\rm t}$, the ratio 
of the true to apparent occurrence rate, taking into account all applicable 
errors.
In Sec.~\ref{sec:model_2} we allow for a distribution of the mass ratio $dn/dq 
\propto q^\gamma$, 
and a concurrent power-law distribution in the luminosity ratio, $dn/dL 
\propto L^\beta$. 
We give expressions for $\Gamma_{\rm a}/\Gamma_{\rm t}$, and evaluate them 
numerically for the case of planets orbiting Sun-like dwarf stars.
%We assume the secondary is not of the desired stellar type.
%NOTE: LB notes that yes, while analytically simple, it doesn't make sense. 
%For the twin binaries case, what possible reason could there be for the 
%secondary to not be of the desired stellar type? Similar for the high mass 
%ratio binaries -- the stars are ~identical.
In Sec.~\ref{sec:model_3}, 
we assume the planets have a power-law radius distribution, ${\rm 
d}N_p/{\rm d}R_p\propto R_p^\alpha$.  We calculate the 
effects of binarity on the apparent $\Gamma_{\rm a}(R_p)$ taking into account 
all relevant errors.
Accounting only for dilution \& its numerator effect, big radius planets have
underestimated inferred rates, the smallest radius planets have overestimated
inferred rates, and the in-between radius planets are in-between.
One consequence is that the claimed HJ occurrence rate discrepancy between 
radial velocity \& transit surveys gets much weaker when accounting for 
binarity.
In Sec.~\ref{sec:discussion} we summarize these models and discuss their 
utility in interpreting transit surveys.

@@

Assume this astronomer lives in a universe in which the following assumptions 
hold.
\begin{itemize}
	\item The true population of ``points'' (stellar systems, all unresolved) 
	comprises only single and double star systems. Single star systems have 
	luminosity in the observed bandpass $L_1$, radii $R_1$, and effective 
	temperature $T_{\rm eff,1}$.
	Double star systems have luminosity in the observed bandpass $L_d = 
	(1+\gamma_R)L_1$, for $\gamma_R = L_2/L_1$ the ratio of the luminosity of 
	the secondary to the primary. 
	In this section, $\gamma_R=1$ across the population 
	of star systems~--~all binaries are assumed to be twins.
	The ratio of the number density of binary systems and the number density of 
	all systems in a volume-limited sample is the binary fraction\footnote{The 
	binary fraction is equivalent to the multiplicity fraction if there are 
	no triple, quadruple, $\ldots$ systems.}.
	\item The true population of planets around these stars is as follows:
	\begin{itemize}
	\item A fraction $\Gamma_{t,s}$ of stars in single star systems 
	have planets of radius $R_p$, with orbital period $P$.
	\item A fraction $\Gamma_{t,d}$ of primary stars in double star 
	systems have planets of radius $R_p$, with orbital period $P$. 
	A fraction $w_{d2}\Gamma_{t,d}$ of secondary stars in a double star 
	systems have the same type of 
	planet, where the weight factor $w_{d2}$ specifies the secondary star 
	occurrence rate as a fraction	of the primary star occurrence rate.
	Any astrophysical difference in planet formation between singles and 
	binaries is captured by $\Gamma_{t,s}, \Gamma_{t,d},$ and $w_{d2}$.
	If one assumes the primary and secondary of a binary system host planets at 
	equal rates, this is the $w_{d2}=1$ limit.
	Similarly, if one assumes  secondaries 
	cannot host planets, this is the $w_{d2}=0$ limit.
	\end{itemize}
\end{itemize}

@@

However, the distance out to which double stars are {\it selected} is 
different from the distance $d_{{\rm max},d}^{\rm p}$ out to which planets can 
be {\it detected} (around the primary of a double star system).
For the binary twin population, since ${\rm S/N} \propto \mathcal{D} 
L^{1/2} d^{-1}$,
\begin{equation}
d_{{\rm max},d}^{\rm p} = (1 + \gamma_R)^{-1/2} d_{{\rm max},s} = 
(1 + \gamma_R)^{-1} d_{{\rm max},d}.
\end{equation}
For $\gamma_R = 1$, this means only 1 in 8 apparently-searchable stars which 
are in fact binary systems can yield detectable planets (the ratio of 
volume in which planets can be detected to the actual selected volume).

@@

To develop more analytic expressions in terms of observables, assume that the 
stars are homogeneously distributed\footnote{
If we wished to write a stellar number density profile that accounted for the 
vertical structure of the Milky Way, we might choose a profile either $\propto 
\exp(-z/H)$, or $\propto {\rm sech}^2(z/H)$ for $z$ the distance from the 
galactic midplane and $H$ a scale-height. Both density profiles would lead 
closed form analytic solutions.	}.
Counting the number of single and double systems in their respective selected 
volumes,
\begin{equation}
N_i = n_{i} \frac{4\pi}{3} d_{{\rm max},i}^3,
\label{eq:number_systems}
\end{equation}
for $i\in{\rm \{single,double\} } \equiv { \{s,d\} }$, and
\begin{equation}
\frac{n_{d}}{n_s+n_d} = {\rm binary\ fraction} \equiv {\rm BF}
\end{equation}
The absolute normalization of the number density $n_s + n_d$ is an observed
quantity [Bovy 2017], as is the binary fraction.
For ``sun-like'' dwarfs with $0.7<M_\star/M_\odot<1.3$,
${\rm BF}=0.44\pm0.02$ [Duchene \& Kraus, 2013, Raghavan et al 2010]. 
Note that these latter authors really quote a {\it multiplicity} fraction of
0.44 for sun-like dwarfs~--~for simplicity we count the $\approx 25\%$ of 
multiple systems that are higher order multiples as doubles.

$d_{{\rm max},i}$ in Eq.~\ref{eq:number_systems} is the maximum distance 
corresponding to the given magnitude limit:
\begin{equation}
d_{{\rm max},i} = \left(\frac{L_{i}}{4 \pi F_{{\rm min}}}\right)^{1/2},
\label{eq:d_max}
\end{equation}
where the limiting flux in the bandpass $F_{{\rm min}}$ could equivalently be 
stated in terms of a limiting magnitude $m_{{\rm min}}$.
In Eq.~\ref{eq:d_max}, again $i\in{\rm \{single,double\} }$. As a 
consequence of imposing a magnitude cut, the maximum distance to which binary 
stars will be selected is greater than that of single stars.
The ratio of double to single systems in the observed SNR-limited sample is 
then
\begin{align}
\frac{N_d}{N_s} &= 
	\frac{n_d}{n_s}
	\left( \frac{d_{{\rm max}, d} }{d_{{\rm max}, s}} \right)^3 \\
&= \frac{{\rm BF}}{1 - {\rm BF}} \times (1+\gamma_R)^{3/2}.
\label{eq:beta_model1}
\end{align}
In the case of twin binaries ($\gamma_R = 1$), with a binary fraction 
${\rm BF} = 0.5$, there are 
$2^{3/2}\approx2.8$ times more binary systems than single systems in the 
SNR-limited sample.
Since the completeness fraction for binary systems is $1/8$, there are 
$2^{-3/2}\approx0.35$ times fewer binary systems in the sample that can yield 
planets as single star systems in the system that can yield planets.

@@

One parametrization for the distribution of 
binary mass ratios in a volume-limited sample is $f(q)\propto q^\gamma$.
[Duchene and Kraus 2013], fitting all the multiple systems of [Raghavan et al 
(2010)]'s Fig 16, quote $\gamma = 0.28\pm0.05$ for $0.7<M_\star/M_\odot<1.3$.
Examining only the binary systems of Rhagavan et al 2010, Fig 16, the 
distribution is roughly uniform,
\begin{equation}
{\rm prob}(q) =
\Bigg\{\begin{array}{lr}
c_q & 0.1 < q \leq 1  \\
0 & {\rm otherwise},
\end{array}
\label{eq:mass_ratio}
\end{equation}
provided we ignore the peak at high mass ratios for analytic convenience. $c_q 
= 1/9$ is the normalization.
We choose this latter mass ratio distribution.

For the mass-luminosity relation, for analytic convenience we assume 
$L \propto M^\alpha$, with the lore-value of $\alpha$ being $3.5$.
As a simple check, we fit lines to measured masses and luminosities of dwarf 
stars\footnote{
    Data collected by Torres et al. [2009] for dwarfs earlier than 
    spectral class M, and Benedict et al. [2016] for dwarfs later than 
    spectral class M.
    To convert Benedict et al. [2016]'s reported $M_V$ values to absolute 
    luminosities, we interpolated over E. Mamajek's 
    table
	\texttt{pas.rochester.edu/\textasciitilde 
		emamajek/EEM\_dwarf\_UBVIJHK\_colors\_Teff.txt},
	downloaded 2017.08.02}.
In log-log space, we let the intersection point of the lines float. The 
resulting fit parameters are available in a footnote\footnote{
	m lo: 1.8818719873988132~--~
	c lo: -0.9799647314108376~--~
	m hi: 5.1540712426599882~--~
	c hi: 0.0127626185389781~--~
	M at merge: 0.4972991257826812~--~
	L at merge: 0.0281260412126928.
}. Though the data show a break at $M_\star \approx 0.5M_\odot$, the 
$L\propto M^{3.5}$ relation is a reasonable fit for $M\lesssim 2M_\odot$.

The final addition to our model is that we must assume a stellar mass-radius 
relation.
This affects the transit probabilities and signal sizes about the 
secondaries of binary star systems.
Since again, our main purpose is to develop intuition about observational 
biases, we coarsely assume
\begin{equation}
R\propto M.
\end{equation}
\begin{comment}
[Demircan \& Kahraman 1991]'s empirical fit to eclipsing 
binary data,
\begin{equation}
R = 1.06 M^{0.945}
\label{eq:mass_radius}
\end{equation}
applies for all the stars in our desired mass range ($M < 1.66M_\odot$, D\&K 
1991's stated bounds), where each quantity is given in solar units. 
\end{comment}
With our model specified, we ask:

>>>>

An astronomer who does not believe in binaries wants to measure the
occurrence of planets of a certain type around stars of a certain type.
In other words, they wish to find the mean number of planets within specific 
planetary bounds orbiting the stars in a given volume 
of stellar phase-space $(\Delta R_p, \Delta P, \Delta M_\star,\ldots$).
They perform a signal-to-noise limited transit survey and detect $N_{\rm det}$ 
transit signals that appear to be from planets of the desired type.  They 
calculate the number of stars $N_\star$ that seem to be of the desired type, 
and searchable for the planet.  These ``searchable stars'' are the points on 
the sky for which they think planets are observed with 100\% detection 
efficiency. Correcting for the geometric 
transit probability $f_{\rm g}$, they compute an apparent occurrence rate 
$\Gamma_{\rm a}$:

@@

Here we focus on problems that arise from the fact that stars of the desired
type often exist in binaries. 
We assume for simplicity that all binaries in the survey are 
spatially unresolved.

An immediate complication is that~--~due to dynamical stability or some 
aspect of planet formation~--~ the occurrence rate of planets may differ 
between binary and single-star systems.
If ``occurrence rate'' is defined purely as the mean number of planets within 
set radius and period bounds per star in a given interval of say $\Delta 
M_\star$, it must implicitly marginalize over stellar multiplicity.
Otherwise, one must address ``occurrence rates in single star systems'', 
``occurrence rates about primaries of double systems'', and
``occurrence rates about secondaries of double star systems''.

Outside of astrophysical differences, there are observational biases.
A given apparently-searchable star may truly be a single star of the desired 
type. If not, and ignoring higher order multiples, one of the following must 
be true:
\begin{enumerate}	
\item The apparently-searchable star is a binary system, with two searchable 
stars of the desired type. $N_\star$ is under-counted by one for every such 
system.
%
\item The apparently-searchable star is a binary system, with one searchable 
star of the desired type. 
The searchable star of the desired type could be the primary or the secondary.
$N_\star$ is correctly counted for every such system, though there may be 
systematic errors in estimates of stellar parameters.
%
\item  The apparently-searchable star is a binary system, in which neither of 
the components is a searchable star, but at least one is of the desired type. 
$N_\star$ is over-counted by one for every such system.
\end{enumerate}

@@

When counting $N_{\rm det}$, the detected signals from planets of the desired 
type, neglecting binarity introduces the following error cases:
\begin{enumerate}
\item 
A detected signal is from a star of undesired type (by the above assumptions, 
the secondary of a binary system), from a planet of undesired type, and is 
incorrectly counted as a planet of desired type.
%
\item
A detected signal is from a star of desired type (the primary or secondary of 
a binary), from a planet of undesired type, and is incorrectly counted 
as a planet of desired type.
This and the case above occur when the host star parameters are miscalculated 
({\it e.g.,} the host star is the faint secondary, but is assumed to be 
primary), or when the constant diluting light from the binary companion is not 
corrected.
$N_{\rm det}$ should be lowered by 1 for every such case.
%
\item
A detected signal is from a star of desired type, from a planet of desired 
type, but incorrectly counted as planet of undesired type.
$N_{\rm det}$ should be raised by 1 for every such case.
\item
An {\it undetected} signal from a planet of the desired type, around a star of 
the desired type, was not detected because of dilution from the companion 
(because the SNR floor of the survey is set for single stars). 
This happens during case \#3 in $N_\star$-counting errors (the 
apparently-searchable star is a binary system, in which neither component is 
in fact searchable). If $N_\star$ is corrected for this, {\it i.e.,} the 
system is correctly counted as ``not searchable'', no correction to $N_{\rm 
det}$ is needed.
\end{enumerate}

@@

Our approach is step-by-step starting from a very simple scenario.
In Sec.~\ref{sec:model_1} we first consider a universe consisting of only the 
desired types of stars and planets.  
We give an analytic expression for $\Gamma_{\rm a}/\Gamma_{\rm t}$, the ratio 
of the true to apparent occurrence rate, taking into account all applicable 
errors.
In Sec.~\ref{sec:model_2} {\bf we allow for a distribution of the mass ratio 
$dn/dq \propto q^\gamma$} {\bf (not done yet)}, 
and a concurrent power-law distribution in the luminosity ratio, $dn/dL 
\propto L^\beta$. 
We give expressions for $\Gamma_{\rm a}/\Gamma_{\rm t}$, and evaluate them 
numerically for the case of planets orbiting Sun-like dwarf stars.
%We assume the secondary is not of the desired stellar type.
%NOTE: LB notes that yes, while analytically simple, it doesn't make sense. 
%For the twin binaries case, what possible reason could there be for the 
%secondary to not be of the desired stellar type? Similar for the high mass 
%ratio binaries -- the stars are ~identical.
In Sec.~\ref{sec:model_3}, 
we assume the planets have a power-law radius distribution, ${\rm 
d}N_p/{\rm d}R_p\propto R_p^\alpha$.  We calculate the 
effects of binarity on the apparent $\Gamma_{\rm a}(R_p)$ taking into account 
all relevant errors.
Accounting only for dilution \& its numerator effect, big radius planets have
underestimated inferred rates, the smallest radius planets have overestimated
inferred rates, and the in-between radius planets are in-between.
One consequence is that the claimed HJ occurrence rate discrepancy between 
radial velocity \& transit surveys gets much weaker when accounting for 
binarity.
In Sec.~\ref{sec:discussion} we summarize the models and discuss how they can 
be used to interpret transit surveys.

@@

Assume this astronomer lives in a universe in which the following assumptions 
hold.
\begin{itemize}
	\item The true population of ``points'' (stellar systems, all unresolved) 
	comprises only single and double star systems. Single star systems have 
	luminosity in the observed bandpass $L_1$, radii $R_1$, and effective 
	temperature $T_{\rm eff,1}$.
	Double star systems have luminosity in the observed bandpass $L_d = 
	(1+\gamma_R)L_1$, for $\gamma_R = L_2/L_1$ the ratio of the luminosity of 
	the secondary to the primary. 
	In this section, $\gamma_R=1$ across the population 
	of star systems~--~all binaries are assumed to be twins.
	The ratio of the number density of binary systems to all systems in a 
	volume-limited sample is the binary fraction\footnote{The 
	binary fraction is equivalent to the multiplicity fraction if there are 
	no triple, quadruple, $\ldots$ systems.}.
	\item The true population of planets around these stars is as follows.
    In single star systems, a fraction $\Gamma_{t,s}$ of stars 
	have planets of radius $R_p$, with orbital period $P$.
	In double star systems, a fraction $\Gamma_{t,d}$ of primary stars have 
	planets of radius $R_p$, with orbital period $P$. 
	A fraction $w_{d2}\Gamma_{t,d}$ of secondary stars in double star 
	systems have the same type of 
	planet, where the weight factor $w_{d2}$ specifies the secondary star 
	occurrence rate as a fraction	of the primary star occurrence rate.
	Any astrophysical difference in planet formation between singles and 
	binaries is captured by $\Gamma_{t,s}, \Gamma_{t,d},$ and $w_{d2}$.
	If one assumes the primary and secondary of a binary system host planets 
	at equal rates, this is the $w_{d2}=1$ limit.
	Similarly, if one assumes  secondaries 
	cannot host planets, this is the $w_{d2}=0$ limit.
\end{itemize}

@@

However, the distance out to which double stars are {\it selected} is 
different from the distance $d_{{\rm max},d}^{\rm p}$ out to which planets can 
be {\it detected} (around the primary of a double star system).
For the binary twin population, all planet and stellar radii are identical, so 
${\rm S/N} \propto \mathcal{D}, and
L^{1/2} d^{-1}$,
\begin{equation}
d_{{\rm max},d}^{\rm p} = (1 + \gamma_R)^{-1/2} d_{{\rm max},s} = 
(1 + \gamma_R)^{-1} d_{{\rm max},d}.
\end{equation}
For $\gamma_R = 1$, this means only 1 in 8 apparently-searchable stars which 
are in fact binary systems can yield detectable planets (the ratio of 
volume in which planets can be detected to the actual selected volume).

@@

To develop more analytic expressions in terms of observables, assume that the 
stars are homogeneously distributed\footnote{
If we wished to write a stellar number density profile that accounted for the 
vertical structure of the Milky Way, we might choose a profile either $\propto 
\exp(-z/H)$, or $\propto {\rm sech}^2(z/H)$ for $z$ the distance from the 
galactic midplane and $H$ a scale-height. Both density profiles would lead 
closed form analytic solutions.	}.
Counting the number of single and double systems in their respective selected 
volumes,
\begin{equation}
N_i = n_{i} \frac{4\pi}{3} d_{{\rm max},i}^3,
\label{eq:number_systems}
\end{equation}
for $i\in{\rm \{single,double\} } \equiv { \{s,d\} }$, and
\begin{equation}
\frac{n_{d}}{n_s+n_d} = {\rm binary\ fraction} \equiv {\rm BF}
\end{equation}
The absolute normalization of the number density $n_s + n_d$ is an observed
quantity [Bovy 2017], as is the binary fraction.
For ``sun-like'' dwarfs with $0.7<M_\star/M_\odot<1.3$,
${\rm BF}=0.44\pm0.02$ [Duchene \& Kraus, 2013, Raghavan et al 2010]. 
These latter authors really quote a {\it multiplicity} fraction of
0.44 for sun-like dwarfs~--~for simplicity we count the $\approx 25\%$ of 
multiple systems that are higher order multiples as doubles.

$d_{{\rm max},i}$ in Eq.~\ref{eq:number_systems} is the maximum distance 
corresponding to the given magnitude limit:
\begin{equation}
d_{{\rm max},i} = \left(\frac{L_{i}}{4 \pi F_{{\rm min}}}\right)^{1/2},
\label{eq:d_max}
\end{equation}
where the limiting flux in the bandpass $F_{{\rm min}}$ could equivalently be 
stated in terms of a limiting magnitude $m_{{\rm min}}$.
In Eq.~\ref{eq:d_max}, again $i\in{\rm \{single,double\} }$. As a 
consequence of imposing a magnitude cut, the maximum distance to which binary 
stars will be selected is greater than that of single stars.
The ratio of double to single systems in the observed sample is 
then
\begin{align}
\frac{N_d}{N_s} &= 
	\frac{n_d}{n_s}
	\left( \frac{d_{{\rm max}, d} }{d_{{\rm max}, s}} \right)^3 \\
&= \frac{{\rm BF}}{1 - {\rm BF}} \times (1+\gamma_R)^{3/2}.
\label{eq:beta_model1}
\end{align}
In the case of twin binaries ($\gamma_R = 1$), with a binary fraction 
${\rm BF} = 0.5$, there are 
$2^{3/2}\approx2.8$ times more binary systems than single systems in the 
SNR-limited sample.
Since the completeness fraction for binary systems is $1/8$, there are 
$2^{-3/2}\approx0.35$ times fewer binary systems in the sample that can yield 
planets as single star systems in the system that can yield planets.

@@

One parametrization for the distribution of 
binary mass ratios in a volume-limited sample is $f(q)\propto q^\gamma$.
[Duchene and Kraus 2013], fitting all the multiple systems of [Raghavan et al 
(2010)]'s Fig 16, quote $\gamma = 0.28\pm0.05$ for $0.7<M_\star/M_\odot<1.3$.
Examining only the binary systems of Rhagavan et al 2010, Fig 16, the 
distribution is roughly uniform,
\begin{equation}
{\rm prob}(q) =
\Bigg\{\begin{array}{lr}
{\rm constant} & 0.1 < q \leq 1  \\
0 & {\rm otherwise},
\end{array}
\label{eq:mass_ratio}
\end{equation}
though we ignore the peak at high mass ratios for analytic convenience. 
We choose this latter mass ratio distribution.

For the mass-luminosity relation, %for analytic convenience 
we assume 
$L \propto M^\alpha$, with the lore-value of $\alpha$ being $3.5$.
To check, we fit lines to measured masses and luminosities of dwarf 
stars\footnote{
    Data collected by Torres et al. [2009] for dwarfs earlier than 
    spectral class M, and Benedict et al. [2016] for dwarfs later than 
    spectral class M.
    To convert Benedict et al. [2016]'s reported $M_V$ values to absolute 
    luminosities, we interpolated over E. Mamajek's 
    table
	\texttt{pas.rochester.edu/\textasciitilde 
		emamajek/EEM\_dwarf\_UBVIJHK\_colors\_Teff.txt},
	downloaded 2017.08.02}.
In log-log space, we let the intersection point of the lines float. The 
resulting fit parameters are available in a footnote\footnote{
	m lo: 1.8818719873988132~--~
	c lo: -0.9799647314108376~--~
	m hi: 5.1540712426599882~--~
	c hi: 0.0127626185389781~--~
	M at merge: 0.4972991257826812~--~
	L at merge: 0.0281260412126928.
}. Though the data show a break at $M_\star \approx 0.5M_\odot$, the 
$L\propto M^{3.5}$ relation is a reasonable fit for $M\lesssim 2M_\odot$.

The final necessary assumption is a stellar mass-radius relation.
This affects the transit probabilities and signal sizes about the 
secondaries of binary star systems.
Since our main purpose is to develop intuition about observational 
biases, we coarsely assume
\begin{equation}
R\propto M.
\end{equation}
\begin{comment}
[Demircan \& Kahraman 1991]'s empirical fit to eclipsing 
binary data,
\begin{equation}
R = 1.06 M^{0.945}
\label{eq:mass_radius}
\end{equation}
applies for all the stars in our desired mass range ($M < 1.66M_\odot$, D\&K 
1991's stated bounds), where each quantity is given in solar units. 
\end{comment}
With our model specified, we ask:
